Formula_sensitivity

### 2019-05-28 Analysis/Simulation/Formula_sensitivity.Rmd
Add few temp files to analyze data on road

## Load the libraries we will be using
library(tidyverse)
library(scales)
library(plyr)
library(gridExtra)
```

## Introduction

The growing literature on IRB risk-weights focus on the drivers and consequences of average risk-weights dispersion *across* banks. I intend to show that bank-level risk-weight dispersion, i.e., the standard deviation of risk-weight density *within* a bank, also plays a role. The study has two parts. First, I plan to investigate whether bank-level risk-weight dispersion is associated with better capitalization, and higher return on equity. In the second part, I want to uncover which variables explain the variation of bank-level risk-weight dispersion across banks, i.e., why some banks have a risk-weight distribution wider than others.

## Effects

```{r parameters and data, echo=FALSE, include=FALSE}
set.seed(123)
LGD <- 0.2
M   <- 2.5
PD  <- seq(0.0001, 0.2, by=0.00001) 
PD_1 <- rlnorm(length(PD),-4,0.1)
PD_1[PD_1<0] <- 0.0001
PD_2 <- rlnorm(length(PD),-4,0.5)
PD_2[PD_2<0] <- 0.0001
```

```{r formulas, echo=FALSE, include=FALSE}
mapping <- function(PD,LGD,M){
  R  <- 0.12*(1-exp(-50*PD))/(1-exp(-50))+0.24*(1-(1-exp(-50*PD))/(1-exp(-50)))
  b  <- (0.11852-0.05478*log(PD))^2
  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(0.999))-PD)*((1-1.5*b)^-1*(1+(M-2.5)*b))
  RW <- K*12.5
  data <- data.frame(RW,PD)
  return(data)
}

@@

### Why does risk-weight dispersion matters for capitalization and ROE?

According to the IRB framework, banks can use their own estimates of probability of default (PD).^[This is valid for banks approved both under the Foundation-IRB and Advanced-IRB. Banks under the latter approach may also use their own estimates of loss-given-default (LGD) and exposure-at-defualt (EAD).] Risk-weights are then mapped from these estimated PDs according to a pre-determined formula. The figure 1 shows how estimated PDs map into regulatory risk-weights.^[The figure shows the map for loans in the corporate sector, to large firms, assuming a 20% of loss given defaults and loan maturity of 2.5 years. The choice of assumptions does not affect my reasoning.]

```{r, echo=FALSE}
data.1 <- mapping(PD,LGD,M)

p <- ggplot(data = subset(data.1,subset = PD <1),
            mapping = aes(x = PD, y=RW)) 
p + geom_line(alpha = 1) + 
  labs(x = "PD", 
         y = "RW",title = "Figure 1:") +
  theme_bw()
```

The concativity of the risk-weight function means that, for a given average PD, an increase in the standard deviation of PD reduces the average risk-weight. Conversely, for a given average RW, greater RW dispersion implies higher average PD. Figure 2 illustrate the last point using simulated data.^[PDs are drawn from log-normal distributions and RW computed using the function shown in figure 1.]

```{r,echo=FALSE}
data.1 <- mapping(PD_1,LGD,M)
data.1 <- mutate(data.1,sample=factor("Sample 1"))
data.2 <- mapping(PD_2,LGD,M)
data.2 <- mutate(data.2,sample=factor("Sample 2"))
data <- rbind(data.1, data.2)
mu <- ddply(data, "sample", summarise, rw.mean=mean(RW),pd.mean=mean(PD))
d.mu <- ts(mu)
d.mu <- d.mu/stats::lag(d.mu,-1) - 1

p1 <- ggplot(data = subset(data,subset = PD<0.04),
            mapping = aes(x = PD, fill = sample, color = sample)) + 
  geom_density(alpha = 0.3) +
  geom_vline(data=mu,aes(xintercept = pd.mean,color=sample), linetype="dashed",size=1)+
  annotate(geom = "text", x = 0.021, y = 200,
                            label = "12.5% higher",
                            hjust = 0)+
  labs(x = "PD",
         title = "Figure 2:") +
  theme_classic() + theme(legend.position = "none")

p2 <- ggplot(data = subset(data,subset = RW<0.7 & RW >0.3),
            mapping = aes(x = RW, fill = sample, color = sample)) + 
  geom_density(alpha = 0.3) + 
  geom_vline(data=mu,aes(xintercept = rw.mean,color=sample), linetype="dashed",size=1)+
  annotate(geom = "text", x = 0.53, y = 25,
                            label = "0.3% higher",
                            hjust = 0)+
  labs(x = "RW") +
 theme_classic() + theme(legend.position = "none")

grid.arrange(p1, p2, nrow = 1)

@@

Hence, by widening their RW density, banks can either reduce their risk-weighted asset and maintain their risk-taking level and/or increase risk-taking without affecting their capital ratios.

**Hypothesis 1** *Banks with wider risk-weight distribution are better capitalized*

**Hypothesis 2** *Banks with wider risk-weight distribution have higher return on equity*

# Empirical strategy

The main analysis relies on the following regressions:\newline

$\text{CAR}_{b,t}=\alpha_0+\alpha_1\widehat{sd(\text{RW})}_{b,t}+\alpha_2\text{Controls}_{b,t}+\varepsilon_{b,t}$,  \newline

$\text{ROE}_{b,t}=\beta_0+\beta_1\widehat{sd(\text{RW})}_{b,t}+\beta_2\text{Controls}_{b,t}+\varepsilon_{b,t}$,  \newline

where the independent variables, $\text{CAR}_{b,t}$ and $\text{ROE}_{b,t}$, are the capital ratio and return on equity of bank $b$ at year $t$, respectively. The variable of interest, $\widehat{sd(\text{RW})}_{b,t}$, captures exogenous changes in risk-weight dispersion and it is obtained as the fitted value for the following first stage regression: \newline

$sd(\text{RW})_{b,t}=\kappa_0+\kappa_1\text{IRB}_{b,t}+\kappa_2\text{Controls}_{b,t}+\varepsilon_{b,t}$,  \newline

$sd(\text{RW})_{b,t}$ is the standard deviation of risk-weight of bank $b$ at time $t$. For the years before IRB approval $sd(\text{RW})_{b,t}$ is close to zero.^[Risk-weights under the standardized approach are flat within an asset class.] After approval we replace this variable by the standard deviation of the IRB risk-weight density. The instrument, $IRB_{b,t}$, is a dummy variable that takes 1 for the periods after bank $b$ is allowed to calculate risk-weights using IRB models.

## Causes and policy 
After stabilishing the importance of risk-weight dispersion, I focus on the policy implications of this finding.

$\text{sd(RW)}_{b,c,t}=\gamma_0+\gamma_1\text{Regulation}_{c,t}+\gamma_2\text{Bank}_{b,t}+\varepsilon_{b,c,t}$,  \newline

$\text{Regulation}_{c,t}$ is a vector of variables that may include regulatory standards, for instance, from the Barth dataset on bank regulation. $\text{Bank}_{b,t}$ is a vector bank characteristics that may affect the dispersion of risk-weight, such as, NPL ratio, size and diversification. Comparing the economic and statistical significance between $\gamma_1$ and $\gamma_2$ shows to what extent risk-weight dispersion is the unintended result of regulation or specific to banks' strategies and possibilities.\newline

## Data

For this project, I will have to merge the Bankscope dataset to risk-weight information from Pillar-III reports, which we are already collecting.

>>>>

## Load the libraries we will be using
library(tidyverse)
library(scales)
library(plyr)
library(gridExtra)
library(DescTools)
library(lfe)
library(stargazer)
```

```{r parameters and data, echo=FALSE, include=FALSE}
set.seed(123)
LGD <- 0.2
M   <- 2.5
grid  <- seq(0.0001, 1, by=0.00001) 
grid_2  <- seq(0, 10, by=0.01) 
PD_1 <- rlnorm(length(grid),-4,0.1)
PD_1[PD_1<0] <- 0.0001
PD_1_avg <- mean(PD_1)
PD_1_sd <- sd(PD_1)
PD_2 <- rlnorm(length(grid),-4,0.5)
PD_2[PD_2<0] <- 0.0001
PD_2[PD_2>0.9999] <- 0.9999
PD_2_avg <- mean(PD_2)
PD_2_sd <- sd(PD_2)
```

```{r formulas, echo=FALSE, include=FALSE}
mapping_wholesale <- function(PD,LGD,M,alpha,portfolio){
  R  <- 0.12*(1-exp(-50*PD))/(1-exp(-50))+0.24*(1-(1-exp(-50*PD))/(1-exp(-50)))
  b  <- (0.11852-0.05478*log(PD))^2
  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)*((1-1.5*b)^-1*(1+(M-2.5)*b))
  RW <- K*12.5
  data <- data.frame(K,RW,PD,LGD,M,R,portfolio)
  return(data)
}

mapping_retail <- function(PD,LGD,R,alpha,portfolio){
  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
  RW <- K*12.5
  data <- data.frame(K,RW,PD,LGD,M,R,portfolio)
  return(data)
}

mapping_other_retail <- function(PD,LGD,alpha,portfolio){
  R  <- 0.03*(1-exp(-35*PD))/(1-exp(-35))+0.16*(1-(1-exp(-35*PD))/(1-exp(-35)))
  K  <- LGD*(pnorm(((1-R)^(-0.5))*qnorm(PD)+((R/(1-R))^(0.5))*qnorm(alpha))-PD)
  RW <- K*12.5
  data <- data.frame(K,RW,PD,LGD,R,portfolio)
  return(data)
}
```

```{r wholesale vs retail, echo=FALSE}
data <- mapping_wholesale(grid,0.45,2.5,0.999,factor("Corporate"))
data <- rbind(data,mapping_retail(grid,0.45,0.15,0.999,factor("Mortgage")))

p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K, colour = portfolio)) 
p + geom_line(alpha = 1) + 
  labs(x = "PD", 
         y = "Capital requirement per unit of exposure",title = "Figure 1: Wholesale vs Retail") +
  theme_bw()
```

```{r LGD sensitivity, echo=FALSE}
data <- mapping_wholesale(grid,0.4,2.5,0.999,factor("LGD = 40%"))
data <- rbind(data,mapping_wholesale(grid,0.45,2.5,0.999,factor("LGD = 45%")))
data <- rbind(data,mapping_wholesale(grid,0.5,2.5,0.999,factor("LGD = 50%")))

p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K, colour = portfolio)) 
p + geom_line(alpha = 1) + 
  labs(x = "PD", 
         y = "Capital requirement per unit of exposure",title = "Figure 2: PD mapping") +
  theme_bw()
```
```{r M sensitivity, echo=FALSE}
data <- mapping_wholesale(grid,0.1,0,0.999,factor("M = 0 year"))
data <- rbind(data,mapping_wholesale(grid,0.1,2.5,0.999,factor("M = 2.5 years")))
data <- rbind(data,mapping_wholesale(grid,0.1,5,0.999,factor("M = 5 years")))

p <- ggplot(data = subset(data,subset = PD <0.2),
            mapping = aes(x = PD, y=K, colour = portfolio)) 
p + geom_line(alpha = 1) + 
  labs(x = "PD", 
         y = "Capital requirement per unit of exposure",title = "Figure 3: PD mapping") +
  theme_bw()
```

```{r LGD_mapping, echo=FALSE}
data <- mapping_wholesale(0.01,grid,2.5,0.999,factor("PD = 1%"))
data <- rbind(data,mapping_wholesale(0.05,grid,2.5,0.999,factor("PD = 5%")))
data <- rbind(data,mapping_wholesale(0.1,grid,2.5,0.999,factor("PD = 10%")))

p <- ggplot(data = subset(data,subset = PD <1),
            mapping = aes(x = LGD, y=K, colour = portfolio)) 
p + geom_line(alpha = 1) + 
  labs(x = "LGD", 
         y = "Capital requirement per unit of exposure",title = "Figure 4: LGD mapping") +
  theme_bw()
```

```{r M_mapping, echo=FALSE}
data <- mapping_wholesale(0.01,0.45,grid_2,0.999,factor("PD = 1%"))
data <- rbind(data,mapping_wholesale(0.05,0.45,grid_2,0.999,factor("PD = 5%")))
data <- rbind(data,mapping_wholesale(0.1,0.45,grid_2,0.999,factor("PD = 10%")))

p <- ggplot(data = data,
            mapping = aes(x = M, y=K, colour = portfolio)) 
p + geom_line(alpha = 1) + 
  labs(x = "Maturity", 
         y = "Capital requirement per unit of exposure",title = "Figure 5: Maturity mapping") +
  theme_bw()
```

```{r savings simulation, echo=FALSE,include=FALSE}

K_bar <- 99991*mapping_wholesale(mean(PD_1),0.45,2.5,0.999,factor("Average PD"))["K"]
K_i <- mapping_wholesale(PD_1,0.45,2.5,0.999,factor("PD distribution"))
new <- summarise(K_i,EAD = Freq(K))
K_total <- K_i  %>%
  group_by(portfolio) %>% 
  summarise_all(sum)

print(K_bar-K_total["K"])
print(K_total["K"]/K_bar-1)

K_bar <- 99991*mapping_wholesale(mean(PD_2),0.45,2.5,0.999,factor("Average PD"))["K"]
K_i <- mapping_wholesale(PD_2,0.45,2.5,0.999,factor("PD distribution"))
K_total <- K_i  %>%
  group_by(portfolio) %>% 
  summarise_all(sum)

print(K_bar-K_total["K"])
print(K_total["K"]/K_bar-1)

K_bar <- 99991*mapping_wholesale(mean(PD_2),0.55,2.5,0.999,factor("Average PD"))["K"]
K_i <- mapping_wholesale(PD_2,0.55,2.5,0.999,factor("PD distribution"))
K_total <- K_i  %>%
  group_by(portfolio) %>% 
  summarise_all(sum)

print(K_bar-K_total["K"])
print(K_total["K"]/K_bar-1)

K_bar <- 99991*mapping_wholesale(mean(PD_2),0.45,0,0.999,factor("Average PD"))["K"]
K_i <- mapping_wholesale(PD_2,0.45,0,0.999,factor("PD distribution"))
K_total <- K_i  %>%
  group_by(portfolio) %>% 
  summarise_all(sum)

print(K_bar-K_total["K"])
print(K_total["K"]/K_bar-1)

@@

```{r PD LGD correlation, echo=FALSE}
load("../../Data/Temp/PD_LGD_correlation.Rda")

reg.hdfe1 <- felm(LGD~ PD | Portfolio+Bank|0|0,
                  data = subset(data,subset = PD<1))
robust.hdfe1 <- summary(reg.hdfe1,robust=TRUE)$coefficients[,2]

reg.hdfe2 <- felm(LGD~ PD | Bank|0|0,
                  data = subset(data,subset = PD<1 & Portfolio == "Real estate"))
robust.hdfe2 <- summary(reg.hdfe2,robust=TRUE)$coefficients[,2]

reg.hdfe3 <- felm(LGD~ PD | Bank|0|0,
                  data = subset(data,subset = PD<1 & Portfolio == "Qualifying revolving" ))
robust.hdfe3 <- summary(reg.hdfe3,robust=TRUE)$coefficients[,2]

reg.hdfe4 <- felm(LGD~ PD |  Bank|0|0,
                  data = subset(data,subset = PD<1 & Portfolio == "Other retail"))
robust.hdfe4 <- summary(reg.hdfe4,robust=TRUE)$coefficients[,2]

reg.hdfe5 <- felm(LGD~ PD |  Bank|0|0,
                  data = subset(data,subset = PD<1 & Portfolio == "Corporate"))
robust.hdfe5 <- summary(reg.hdfe5,robust=TRUE)$coefficients[,2]

reg.hdfe6 <- felm(LGD~ PD |  Bank|0|0,
                  data = subset(data,subset = PD<1 & Portfolio == "Banks"))
robust.hdfe6 <- summary(reg.hdfe6,robust=TRUE)$coefficients[,2]

reg.hdfe7 <- felm(LGD~ PD |  Bank|0|0,
                  data = subset(data,subset = PD<1 & Portfolio == "Sovereign"))
robust.hdfe7 <- summary(reg.hdfe7,robust=TRUE)$coefficients[,2]
```
```{r PD LGD table, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.hdfe1,reg.hdfe2,reg.hdfe3,reg.hdfe4,reg.hdfe5,reg.hdfe6,reg.hdfe7,
          header = FALSE, type="latex", keep.stat = c("n","rsq"), font.size = "small", 
          label = "",
          style = "qje", 
          column.labels   = c("All", "Real estate","Credit cards", "Other retail","Corporates","Banks","Sovereign"),
          dep.var.labels = c("LGD"), column.sep.width = "2pt",
          add.lines = list(c("Portfolio FE", "Yes", "-","-","-","-","-","-"),
                           c("Bank FE","Yes", "Yes","Yes","Yes","Yes","Yes","Yes")))

@@
